Predicting the 2024 US Presidential Election with a Model-Based Forecast

### 2024-10-31 paper/paper.qmd
update scripts, correct intro and abstract, update reference

---
title: "Predicting the 2024 US Presidential Election with a Model-Based Forecast"
subtitle: "Using Generalized Linear Models to Predict Election Outcomes"
author: 
  - Yuanyi (Leo) Liu
  - Dezhen Chen
  - Ziyuan Shen
thanks: "Code and data are available at: [Forecasting the 2024 US Presidential Election](https://github.com/leoyliu/Forecasting-the-2024-US-Presidential-Election)."
date: today
date-format: long
abstract: "In this paper, we have built a linear model to predict the outcome of the 2024 U.S. presidential election, using a \"poll-of-polls\" approach that combines data from many well-known pollsters. Our analysis shows that Donald Trump has a probability of winning in range 45% to 55% in each state, which depends on how polling data changes from state to state. However, Kamala Harris is projected to win key states and receive the majority of electoral votes, with a projected 239 votes to Trump's 98, based on testing dataset, we only have those state in the dataset. These show the value of aggregate polls in providing reliable predictions of election outcomes."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

The results of presidential elections influence national and international policies and determine governance and economic priorities. An accurate election forecast is a valuable tool for political strategists, the media, and the public. The 2024 U.S. presidential election is expected to be highly competitive, with Donald Trump and Kamala Harris as the leading candidates. In this paper, we use national polling data from the period leading up to the 2024 U.S. presidential election to generate a linear regression model that predicts the winner while capturing changes in public opinion throughout the election period.

This study estimates the probability of each candidate winning based on aggregate polling data. By combining the results, we reduce the bias and short-term volatility of individual polls. By using a linear model, we estimate the probability of a Trump victory to be between 45% and 55%, which depends on state-by-state variations.

@@

## Estimand
Todo:

@@

# Results {#sec-result}
@sec-result examines the relationship between polling score, methodology, poll duration, transparency score, sample size, and state, with respect to Donald Trump's polling performance in the 2024 U.S. Presidential Election. Using a dataset that captures polling data across various states and polling organizations, we apply a linear regression model to identify key factors that influence Trump's support. Below, we present the results of our model and its implications.

@@

## Prediction map
@fig-statemap displays the predicted outcome for each state in the 2024 U.S. Presidential Election, with green representing states where Trump is expected to win and blue for those where Harris is predicted to prevail.
```{r}
#| label: fig-statemap
#| fig-cap: Predicted Winner of 2024 US Presidential Election by State
#| echo: false
#| warning: false
#| message: false

# "red" for Trump and "blue" for Harris
state_predictions$winner_color <- ifelse(state_predictions$winner == "Trump", "green", "blue")

# Create a map visualization showing the predicted winner by state
plot_usmap(data = state_predictions, values = "winner_color", regions = "states") +
  scale_fill_manual(values = c("green" = "green", "blue" = "blue"), name = "Predicted Winner", labels = c("Harris", "Trump")) +
  labs(title = "Predicted Winner of the 2024 U.S. Presidential Election by State") +
  theme(legend.position = "right")

```
The map illustrates Trump gaining traction across central and southern states, including Texas and Missouri, which have traditionally leaned Republican. Meanwhile, Harris holds a solid lead in states along the West Coast and in the Northeast, such as California and New York, regions that have a history of Democratic support.

@@

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

@@

## Overview of Siena/NYT
Siena College Research Institute (SCRI) and The New York Times launched a collaborative political poll in July 2013. Siena/NYT  was known for its rigorous methodology and accuracy, which provided its first real-time midterm election poll in 2018.In 2024, FiveThirtyEight named it the most accurate pollster in the United States. The polling methodology emphasizes live phone interviews and representative sampling and focuses on transparency and statistical accuracy to provide a deep understanding of voter sentiment. [@]

## Target Population, Frame, and Sample
* Target Population: Registered voters or likely voters in the United States, especially voters in major battleground states during an election in 2024.
* Sample Frame: U.S. registered voters based on a national voter file maintained by L2, with a cell phone number that matches the voter file.
* Sample Size: Sample sizes vary across different polls. The poll in late October 2024 surveyed over 2,500 voters, and 2,097 of them completed the full survey. The margin of error for likely voters is about ±2.2 percentage points. [@]

## Sample Recruitment
The sample was recruited by telephone. Registered voters were selected from the voter files held by L2 and contacted via their listed telephone numbers including landlines and mobile phone numbers. More than 260,000 calls were made to more than 80,000 voters, with over 98% of respondents using mobile phones. Interviews were conducted by live telephone interviews in both English and Spanish. [@]

## Sampling Approach and Trade-offs
Siena/NYT used a stratified sampling method. Polls were stratified by demographic characteristics such as political party, race and region. This improves accuracy but adds cost and complexity. It also makes it challenging to reach certain populations such as young voters.

@@

### Sample Size
The target sample size was 10,000 respondents, which are stratified by the Stratification Variables described above. This stratification ensures that all major voter groups are proportionately represented. 10,000 samples have a margin of error of about ±1% and a confidence level of 95%, which is appropriate for predicting elections.

>>>>

---
title: "Predicting the 2024 US Presidential Election with a Model-Based Forecast"
subtitle: "Using Generalized Linear Models to Predict Election Outcomes"
author: 
  - Yuanyi (Leo) Liu
  - Dezhen Chen
  - Ziyuan Shen
thanks: "Code and data are available at: [Forecasting the 2024 US Presidential Election](https://github.com/leoyliu/Forecasting-the-2024-US-Presidential-Election)."
date: today
date-format: long
abstract: "In this paper, we develop a linear model to predict the outcome of the 2024 U.S. presidential election using a \"poll-of-polls\" approach that combines data from various prominent pollsters. Our analysis indicates that Donald Trump’s probability of winning varies between 45% and 55% across states, depending on regional polling trends. However, Kamala Harris is projected to secure key states and achieve a majority of electoral votes, with a projected count of 239 to Trump’s 98 based on the testing dataset. These predictions are based on the testing dataset, which lacks data for some states, limiting our ability to produce results for all states. This paper underscore the effectiveness of aggregated polls in producing reliable election forecasts."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

The results of presidential elections influence national and international policies and determine governance and economic priorities. An accurate election forecast is a valuable tool for political strategists, the media, and the public. The 2024 U.S. presidential election is expected to be highly competitive, with Donald Trump and Kamala Harris as the leading candidates. Despite the abundance of polling data, individual polls are often subject to bias and short-term fluctuations, making it difficult to predict election outcomes with confidence. This paper addresses this issue by using aggregate polling data to provide more reliable predictions of election outcomes.

@@

## Estimand

In this study, our estimand is Donald Trump's percentage of polling in each state ahead of the 2024 U.S. presidential election. The object of the estimation is Trump's percentage of polling data in each state based on aggregated information from various sources By using a linear regression model, we aim to capture changes in public opinion over time and provide a clearer understanding of voter preferences and the likely election outcome.

@@

# Results {#sec-result}

@sec-result examines the relationship between polling score, methodology, poll duration, transparency score, sample size, and state, with respect to Donald Trump's polling performance in the 2024 U.S. Presidential Election. Using a dataset that captures polling data across various states and polling organizations, we apply a linear regression model to identify key factors that influence Trump's support. Below, we present the results of our model and its implications.

We predicted the outcome based on our test dataset. However, due to missing data for certain states in our original dataset, we were unable to make predictions for all states. This limitation arises because the states with NA values in the dataset are not represented in the model, preventing accurate forecasts for those regions.

@@

## Prediction map

```{r}
#| label: fig-statemap
#| fig-cap: Predicted Winner of 2024 US Presidential Election by State
#| echo: false
#| warning: false
#| message: false

# "red" for Trump and "blue" for Harris
state_predictions$winner_color <- ifelse(state_predictions$winner == "Trump", "red", "blue")

# Create a map visualization showing the predicted winner by state
plot_usmap(data = state_predictions, values = "winner_color", regions = "states") +
  scale_fill_manual(values = c("red" = "red", "blue" = "blue"), name = "Predicted Winner", labels = c("Harris", "Trump")) +
  labs(title = "Predicted Winner of the 2024 U.S. Presidential Election by State") +
  theme(legend.position = "right")

```
@fig-statemap displays the predicted outcome for each state in the 2024 U.S. Presidential Election, with green representing states where Trump is expected to win and blue for those where Harris is predicted to prevail.

The map illustrates Trump gaining traction across central and southern states, including Texas and Missouri, which have traditionally leaned Republican. Meanwhile, Harris holds a solid lead in states along the West Coast and in the Northeast, such as California and New York, regions that have a history of Democratic support.

@@

## Interpretation of Polling Data and Electoral Implications

This paper examines the factors affecting Donald Trump’s polling performance in the 2024 U.S. Presidential Election. Our linear regression model shows that poll duration and state-specific factors are the strongest predictors of Trump’s support. Longer poll durations tend to produce more stable data by capturing changes in public sentiment over time rather than short-term reactions. The state variable underscores the importance of local dynamics, as individual states bring unique policy priorities and demographic trends to their voter bases. These regional distinctions are essential in the Electoral College system, where state-level support directly impacts the final electoral count.

Transparency and sample size play secondary roles in data reliability. Polls with high transparency scores provide clear information about methodology, making them more reliable for analysis. Larger sample sizes reduce margins of error, adding stability to projections, especially in closely contested states. While both transparency and sample size support reliable predictions, they have less influence on voter preferences than duration and state-specific factors.

In comparison, polling methodology (e.g., live phone, online) and numeric grade (a general reliability rating) have minimal impact on Trump’s support in our model. Although these factors contribute to overall data quality, they do not significantly shape voter sentiment. This finding suggests that regional context and polling consistency over time are more important in understanding Trump’s polling performance in this election than the specific techniques or reputations of polling organizations.

## Influence of State Policies on Voter Preferences

Beyond the mechanics of polling, our analysis suggests that state-specific policies and regional cultural factors could influence voter preferences and ultimately shape electoral outcomes. For instance, certain policies, such as the legalization of recreational marijuana, reflect broader ideological values that may correlate with support for particular candidates. Although our model does not directly analyze the impact of marijuana policies, including these cultural aspects could offer a more comprehensive view of voter alignment.

States with legal recreational marijuana, for example, might lean more progressive or libertarian, potentially aligning with candidates like Harris who advocate for policies seen as liberal or socially progressive. This policy environment might signal broader state values that align with a particular party's stance, beyond just the specific candidate. Conversely, states that prohibit recreational marijuana use may lean more conservative, suggesting they may have stronger support for Trump. Although this relationship is not absolute, understanding such cultural factors helps contextualize the state-level support each candidate may receive and could be valuable in identifying battleground states. Additionally, factors like local economic policies, educational investment, or healthcare reform could influence voter preferences, making these areas ripe for future analysis.

## Limitations of the Data and Model

One limitation of this study is the incomplete dataset, with some states missing from the polling data. This limits our ability to produce a fully accurate prediction, as certain areas are excluded from the model. Additionally, the focus on state-level data may overlook finer voting patterns within states, particularly in districts with unique political identities.

## Future Research Directions

Future studies could improve by incorporating demographic variables such as age, income, and party affiliation, which might offer a clearer understanding of voter preferences. Examining additional state-level policies, like healthcare or education, may also provide insights into factors shaping voter choices.

Another possible enhancement involves increasing the geographic specificity of the model to include congressional districts or smaller regions, which may lead to more precise forecasts, especially in competitive states. As polling data collection methods evolve, combining traditional regression with machine learning techniques could allow for models that adapt to ongoing changes in public opinion, resulting in more timely and adaptable forecasts.

@@

## Overview of Siena/NYT
Siena College Research Institute (SCRI) and The New York Times launched a collaborative political poll in July 2013. Siena/NYT  was known for its rigorous methodology and accuracy, which provided its first real-time midterm election poll in 2018 [@citeNYT]. In 2024, FiveThirtyEight named it the most accurate pollster in the United States. The polling methodology emphasizes live phone interviews and representative sampling and focuses on transparency and statistical accuracy to provide a deep understanding of voter sentiment.

## Target Population, Frame, and Sample
* Target Population: Registered voters or likely voters in the United States, especially voters in major battleground states during an election in 2024.
* Sample Frame: U.S. registered voters based on a national voter file maintained by L2, with a cell phone number that matches the voter file.
* Sample Size: Sample sizes vary across different polls. The poll in late October 2024 surveyed over 2,500 voters, and 2,097 of them completed the full survey. The margin of error for likely voters is about ±2.2 percentage points.

## Sample Recruitment
The sample was recruited by telephone. Registered voters were selected from the voter files held by L2 and contacted via their listed telephone numbers including landlines and mobile phone numbers. More than 260,000 calls were made to more than 80,000 voters, with over 98% of respondents using mobile phones. Interviews were conducted by live telephone interviews in both English and Spanish.

## Sampling Approach and Trade-offs
Siena/NYT used a stratified sampling method [@citeNYT]. Polls were stratified by demographic characteristics such as political party, race and region. This improves accuracy but adds cost and complexity. It also makes it challenging to reach certain populations such as young voters.

@@

### Sample Size
The target sample size was 5,000 respondents, which are stratified by the Stratification Variables described above. This stratification ensures that all major voter groups are proportionately represented. 5,000 samples have a margin of error of about ±1% and a confidence level of 95%, which is appropriate for predicting elections.
